---
title: "Producing Simulations"
author: "Andrea Brizzi"
date: "22/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Call required packages
library(dplyr)
library(ggplot2)
library(reshape2)
library(knitr)

# specify repository path
repo_path <- '~/Documents/mini_project_2/R0_simulations/'
setwd(repo_path)
data_path <- '~/Documents/mini_project_2/R0_simulations/data/'

# load scripts necessary for simulations
source('simulation.R')
source('BP_simulation.R')
```

## Intro
The objective of this Markdown file is to show how to produce simulations according to different models.
Having obtained a 100 simulations for each setting, we then plot them, produce summary statistics and save these.

We include two different types of simulating mechanism:

 * Stochastic SEIR model
 * Branching Process with Gamma Generation Intervals and Poisson distributed offspring

For each of the two methods, we produce simulations with different $R_0$ values, with values $1.5, 2.0, 2.5, 3.0$.
Concerning the SEIR model, we also consider populations of different sizes: $10^4$ and $10^6$.

This is how I produced the simulations:
```{r, eval = FALSE}
# Initialise lists to store results 
sims_SEIR <- list()
sims_BP <- list()

# Specify different scenarios
R0_vec <- c(1.5, 2, 2.5, 3)
pops_vec <- c(10^4, 10^6)

for (R0 in R0_vec){
  
  # prepare names for lists
  name = paste0('R0=', R0)
  
  # simulate Branching Process 
  sims_BP[[name]] <- simulation_BP(N = 100, R0 = R0, seeds = 5)
  
  for (pop in pops_vec){
    
    # prepare name for lists:
    name2 = paste0(name, 'pop=', pop)
    
    # simulate SEIR
    sims_SEIR[[name2]] <- simulate_seir_stochastic(N = 100, arnaught = R0, 
                                          gamma = 1/3, sigma = 1/3.5,
                                          Reporting_fraction = 1,
                                          N0_par = pop,
                                          I0_par = 5)$sims
  }
}

saveRDS(object = sims_BP, file.path(data_path, "sims_BP"))
saveRDS(object = sims_SEIR, file.path(data_path, "sims_SEIR"))
```


## Analyse trajectories{.tabset}
Having saved the simulations, we are now able to load them:
```{r}
sims_BP <- readRDS(file.path(data_path, "sims_BP"))
sims_SEIR <- readRDS(file.path(data_path, "sims_SEIR"))
```
And we can proceed to plot the different trajectories

### Branching Process

Here are the results obtained through a Branching process simulation with:

 * Generation Intervals following a discretised Gamma distribution with mean 6.5 and standard deviation 0.62.
 * Number of secondary infections distributed as a Poisson with mean given by $R_0$.
 
```{r, out.width= "100%"}
data = as.data.frame(sims_BP$`R0=1.5`)
data$time = 1:length(data$sim1)
data.plot <- reshape2:::melt.data.frame(data = data, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
ggplot(data.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Branching Process, R0 = 1.5")


data = as.data.frame(sims_BP$`R0=2`)
data$time = 1:length(data$sim1)
data.plot <- reshape2:::melt.data.frame(data = data, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
ggplot(data.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Branching Process, R0 = 2")

data = as.data.frame(sims_BP$`R0=2.5`)
data$time = 1:length(data$sim1)
data.plot <- reshape2:::melt.data.frame(data = data, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
ggplot(data.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Branching Process, R0 = 2.5")

data = as.data.frame(sims_BP$`R0=3`)
data$time = 1:length(data$sim1)
data.plot <- reshape2:::melt.data.frame(data = data, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
ggplot(data.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Branching Process, R0 = 3")

```

The "waviness" of the trajectories is due to the relatively small variance of the generation Interval, so that almost every generation lasts 6 to 8 days, additionally to the seeds assumed to all be introduced on day 0.
I could probably improve that by seeding 1 infection on the first 6 days.

### Stochastic SEIR

The stochastic SEIR simulations are in a way more realistic, as they assume a finite population in which the pathogen can spread. For this reason, these will present a peak and a slowing down of infections.
```{r, out.width= "100%"}
data1 = sims_SEIR$`R0=1.5pop=10000`
data1$time <- 1:length(data1$sim1)
data1.plot <- reshape2:::melt.data.frame(data = data1, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
data2 = sims_SEIR$`R0=1.5pop=1e+06`
data2$time <- 1:length(data2$sim1)
data2.plot <- reshape2:::melt.data.frame(data = data2, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
gg1 <- ggplot(data1.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Stochastic SEIR: R0 = 1.5, N = 10^4")
gg2 <- ggplot(data2.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Stochastic SEIR: R0 = 1.5, N = 10^6")
gridExtra:::grid.arrange(gg1, gg2, ncol = 2)


data1 = sims_SEIR$`R0=2pop=10000`
data1$time <- 1:length(data1$sim1)
data1.plot <- reshape2:::melt.data.frame(data = data1, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
data2 = sims_SEIR$`R0=2pop=1e+06`
data2$time <- 1:length(data2$sim1)
data2.plot <- reshape2:::melt.data.frame(data = data2, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
gg1 <- ggplot(data1.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Stochastic SEIR: R0 = 2, N = 10^4")
gg2 <- ggplot(data2.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Stochastic SEIR: R0 = 2, N = 10^6")
gridExtra:::grid.arrange(gg1, gg2, ncol = 2)

data1 = sims_SEIR$`R0=2.5pop=10000`
data1$time <- 1:length(data1$sim1)
data1.plot <- reshape2:::melt.data.frame(data = data1, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
data2 = sims_SEIR$`R0=2.5pop=1e+06`
data2$time <- 1:length(data2$sim1)
data2.plot <- reshape2:::melt.data.frame(data = data2, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
gg1 <- ggplot(data1.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Stochastic SEIR: R0 = 2.5, N = 10^4")
gg2 <- ggplot(data2.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Stochastic SEIR: R0 = 2.5, N = 10^6")
gridExtra:::grid.arrange(gg1, gg2, ncol = 2)

data1 = sims_SEIR$`R0=3pop=10000`
data1$time <- 1:length(data1$sim1)
data1.plot <- reshape2:::melt.data.frame(data = data1, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
data2 = sims_SEIR$`R0=3pop=1e+06`
data2$time <- 1:length(data2$sim1)
data2.plot <- reshape2:::melt.data.frame(data = data2, id.vars = "time" ,
                                        variable.name = "simulation", value.name = "incidence")
gg1 <- ggplot(data1.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Stochastic SEIR: R0 = 3, N = 10^4")
gg2 <- ggplot(data2.plot, aes(x = time, y = incidence, group = simulation, color = simulation)) + geom_line() +  theme(legend.position = "none") + ggtitle("Stochastic SEIR: R0 = 3, N = 10^6")
gridExtra:::grid.arrange(gg1, gg2, ncol = 2)

```

The above plots show how the population sizes have a very big impact on the evolution of the incidence even when $R_0$ is identical. 
We conclude by summarizing some key statistics of the different simulations:

* Peak day
* Peak value
* Proportion of died out epidemics

```{r, out.width= "100%"}
data <- sims_SEIR
names <- names(data)
table <- data.frame(name = as.character(), min_peak_day = as.numeric(),
                    max_peak_day = as.numeric(), min_peak = as.numeric(), max_peak = as.numeric(),
                    proportion_flat = as.numeric())
cnames <- colnames(table)

#Function outputting peak info and indicator on flat trajectory
summarise <- function(x){
  peak_value = max(x)
  if (peak_value > 10){ # Maybe define a better way to indicate 'immediate extinction'
    peak_day = which.max(x)[1]
    flat = 0
  }else{
    peak_value = NA
    peak_day = NA
    flat = 1
  }
  return(c(peak_day, peak_value, flat))
}

# Produce table
for (name in names){
  tmp <- data[[name]]
  tmp <- t(apply(tmp, 2, FUN = summarise))
  colnames(tmp) <- c('peak_day', 'peak_value', 'flat_indicator')
  
  # Summarise results 
  tmp <- c(name, range(tmp[,1], na.rm = TRUE), range(tmp[,2], na.rm = TRUE), mean(tmp[,3]))
  table <- rbind(table, tmp)
}
colnames(table) <- cnames
kable(table)

```

