---
title: "Sensitivity Subexponential"
author: "Andrea Brizzi"
date: "25/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# set up directories' paths
repo_path <- '/home/andrea/Documents/mini_project_2/R0_simulations'
data_path <- file.path(repo_path, 'data')
plots_path <- '/home/andrea/Documents/mini_project_2/plots'

# Dependencies
require(EpiEstim)
require(dplyr)
require(ggplot2)
require(png)
require(ggpubr)

# source scripts
source(file.path(repo_path,"scripts/simulation.R"))
source(file.path(repo_path,"scripts/methods.R"))

# generation interval parameters
mean_si <- 6.5
sd_si <- sqrt(3^2 + 3.5^2) # may need to adjust this for Branching Pr

# plots reqs
reqs <- theme(axis.text=element_text(size=9), axis.title=element_text(size=10), legend.text = element_text(size=12))

# simulate
# simulate data
tmp <- file.path(data_path, 'sims_SEIR_subexp.rds')
if(!file.exists(tmp))
{
  alpha = c(0.7,0.75, 0.8, .85, 0.9, 0.95, 1)

  .f <- function(alpha)
  {
    simulate_seir_stochastic(N = 200, arnaught = 3,
                             gamma = 1/3, sigma = 1/3.5,
                             alpha=alpha,
                             reporting_fraction = 1,
                             N0_par = 10^6,I0_par = 5)
  }

sims <- lapply(alpha, .f)
names(sims) <- paste0('alpha=', alpha)

saveRDS(sims, file=file.path(data_path, 'sims_SEIR_subexp.rds'))
}
```

## Maths

The objective is to study the adaptability of our EpiEstim adjustement to simulated data which do not satisfy the exponential growth scenario. 
In particular, we look at the subexponential case described in (Chowell 2016)[https://royalsocietypublishing.org/doi/10.1098/rsif.2016.0659]. 
Here they describe a simple modification of the SIR model to account for population mixing, which consists in modifying the transition rate from comparment S to I.
$$
\frac{dS(t)}{dt} = -\beta S_t\frac{I^\alpha}{N} \\
= -I^{\alpha-1}\beta S_t \frac{I}{N}
$$

We apply the same modification to our stochastic SEIR model, and consider values of $\alpha \in \{0.88, 0.9, 0.92, 0.94\}$, a reporting rate of $\rho=1$ and $\beta_0 = \gamma \cdot 3.$
However, we need to be careful with the definition of $R_0$. In the 'classical' SEIR, we have: $R_0 =\frac \beta_0 \gamma$, while here the $\beta_0$ is scaled down by a factor of $I_0^{\alpha-1} = 5^{\alpha-1}$. More generally:
$$
R_t = \frac{S_t}{N_t} \frac \beta \gamma I^{\alpha-1} 
$$

## Simulations{.tabset}

### Plots
We can start by simply visualising the curves produced for $R_0=3$
```{r visualisation}
sims <- lapply(c(0.75, 0.8, .85, 0.9, 0.95, 1), function(x){
  tmp <- simulate_seir_stochastic(N=5, gamma=1/3, sigma=1/3.5, alpha=x, I0 = 5, arnaught=3, reporting_fraction = 1, n_t=70)$sims
  tmp <- as.data.frame(t(tmp))
  colnames(tmp) <- 1:ncol(tmp)
  tmp$alpha <- x
  tmp$sim <- gsub('sim','',rownames(tmp))
  tmp <- reshape2::melt(tmp, id.vars = c("alpha", "sim"))
  tmp <- tmp[order(as.numeric(tmp$sim)),]
  colnames(tmp)[c(3,4)] <- c('day', 'incidence')
  tmp
})
sims <- do.call("rbind", sims)
sims <- as.data.frame(sims)
sims$group <- paste0(sims$alpha, '_',sims$sim)
tmp <- sims

gg <- ggplot(data=sims, aes(x=day, y=log(incidence), color=as.factor(alpha), group=group)) + 
  geom_line() +
  scale_x_discrete(breaks=5*1:18) + 
  theme_bw() + theme(legend.position='bottom') + labs(y='logged incidence', color='alpha')
gg
ggsave(filename=file.path(plots_path, 'trajectories_subexponential.tiff'), gg + reqs, device = 'tiff', w=10, h=6, dpi=600)
```


### $R_t$ estimation

Here we compare the evolution of $R_t$ estimates using the different methods for a single simulation.

```{r r_t_comparison}
rho <- 1;trueR0 <- 3; alpha <- .7

set.seed(43)
tmp <- simulate_seir_stochastic(N = 1, arnaught = trueR0,
                                gamma = 1/3, sigma = 1/3.5,
                                alpha=alpha,
                                reporting_fraction = 1,
                                N0_par = 10^6,
                                I0_par = 5,n_t=200)

I <- tmp$I[-length(tmp$I)]
incidence <- tmp$sims[,1]
incidence <- data.frame(cases = incidence, day = seq_along(incidence))
trueRt <- as.numeric(trueR0 * (1-cumsum(incidence$cases/10^6))) * I[-1]^{alpha-1}
trueRt[which(is.infinite(trueRt))] <- NA
trueRt <- data.frame(trueRt=trueRt, day=1:length(trueRt))


# Underreporting & missed generation 
incidence$cases <- rbinom(n = length(incidence$cases),
                              size = incidence$cases,
                              prob = rho)
incidence <- filter(incidence, day >= 14)
first_day <- min(incidence$day)

# == apply normal EpiEstim
WINDOWS <- 7; 
t_start <- 2:(length(incidence$cases) - WINDOWS)
t_end <- t_start + WINDOWS 

epestim <- EpiEstim::estimate_R(c(incidence$cases), method = "parametric_si",
                           config = make_config(
                             list(
                               mean_si = mean_si,
                               std_si = sd_si,
                               t_start = t_start,
                               t_end = t_end
                             )))$R[,c("Mean(R)", "Quantile.0.025(R)", "Quantile.0.975(R)")]

epestim$day <- t_end + first_day - 1 


# == Apply adjusted EpiEstim, using first 2 weeks to infer r.

# log transform cases, add small constant to time-series if zeros are present
if(any(incidence$cases==0)){
  incidence$casesC <- incidence$cases+0.5
  incidence$logcases <- log(incidence$casesC)
}else{
  incidence$logcases <- log(incidence$cases)
}
  
# Fit line through logged cases
exp_fit <- lm(logcases ~ day, data = filter(incidence, day >= first_day & day <= first_day + 13))

# introduce extrapolated data
tmp <- data.frame(day =  first_day - 1 + (-21):0)
tmp$logcases <- predict.lm(exp_fit, newdata = data.frame(day =  first_day - 1 + (-21):0 ))
if ("casesC" %in% names(incidence)){
  tmp$casesC <- exp(tmp$logcases)
  tmp$cases <- tmp$casesC - 0.5
  tmp$cases[tmp$cases < 0] <- 0
}else{
  tmp$cases <- exp(tmp$logcases)
}
# tmp$trueRt <- trueR0

incidence <- rbind(tmp, incidence)

# define windows
START <- which(incidence$day == first_day)
t_start <- START:(length(incidence$cases) - WINDOWS)
t_end <- t_start + WINDOWS 

# add seeds...
epestim_adj <- EpiEstim::estimate_R(c(incidence$cases), method = "parametric_si",
                           config = make_config(
                             list(
                               mean_si = mean_si,
                               std_si = sd_si,
                               t_start = t_start,
                               t_end = t_end
                             )))$R[,c("Mean(R)", "Quantile.0.025(R)", "Quantile.0.975(R)")]

epestim_adj$day <- t_end - START + first_day

# COMPARE
# assign each estimate to the middle of the sliding window:
epestim$day <- epestim$day - WINDOWS/2
epestim_adj$day <- epestim_adj$day - WINDOWS/2

epestim$type = 'Unadjusted'
epestim_adj$type = 'Adjusted'
data <- rbind(epestim, epestim_adj)
colnames(data) <- c('M', 'Low', 'Up', 'day', 'type')
data$type <- gsub('Adjusted','EpEsAdj',data$type)
data$type <- gsub('Unadjusted', 'EpEs', data$type)

gg <- ggplot(data, aes(x = day)) + 
  geom_ribbon(aes(ymin = Low , ymax = Up, color = type, fill = type), alpha = .3 ) +
  geom_line(aes(y = M, color = type), linetype = 'dashed') + 
  geom_line(data=trueRt, aes(x = day, y = trueRt)) + xlim(c(14, 200)) +
  scale_fill_manual(values = c(5,2)) + 
  scale_color_manual(values = c(5,2)) +
  theme_classic2() + theme(legend.position = 'bottom') +
  labs(y= expression("R"[t]), x='day', fill ='', color ='')

rect <- data.frame(xmin=first_day, xmax=first_day + 13, ymin=-Inf, ymax=Inf)
gg + geom_rect(data=rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax),
              fill="grey20",
              alpha=0.5,
              inherit.aes = FALSE)

incidence$before_first <- (incidence$day <= first_day)
incidence$before_first  <- sapply(incidence$before_first, FUN = function(x){if(x){'Imputed'}else{'Observed'}})

gg2 <- ggplot(incidence, aes(x= day, y = logcases, color = before_first )) + 
  geom_point() + 
  geom_abline(slope = exp_fit$coefficients[2], intercept = exp_fit$coefficients[1], color = 'darkgreen') +
  geom_rect(data=rect, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax),
              fill="lightgreen",
              alpha=0.5,
              inherit.aes = FALSE) + 
  theme_classic2() +theme(legend.position = 'bottom') +
  labs(y = "logged cases", color = '' ) 

gg <- ggarrange(gg2, gg,
                 labels="AUTO")

gg
ggsave(filename=file.path(plots_path, 'effects_adjustment_subexponential.tiff'), gg + reqs, device = 'tiff', w=10, h=5, dpi=600)

```

### $R_0$ estimation

We run 100 simulations with $I_0=5$, $\beta_0 = 3\gamma$, reporting rate $\rho=1$.
```{r ro_estimation, warning=FALSE, message = FALSE}

sims <- readRDS(file.path(data_path, 'sims_SEIR_subexp.rds'))
sims <- sims[-1]
rho=1

# Problems showing plots: maybe store them in a list
list_plots <- list()
final_stats <- data.frame(method = as.character(),
                          coverage = as.numeric(),
                          R0 = as.character(),
                          CIwidth = as.numeric(),
                          q_Rt05 = as.numeric(),
                          q_Rt50 = as.numeric(),
                          q_Rt95 = as.numeric()
                          )

names <- names(sims)
for (name in names){
  
  # Transform the data
  df <- as.data.frame(sims[[name]]$sims)
  df$day <- 1:(dim(df)[1])
  
  # Now cut off 2 weeks of data
  df <- filter(df, day > 2*7)
  
  results <- data.frame(sim = as.numeric(),
                        method = as.character(),
                        R0 = as.numeric(),
                        lowR0 = as.numeric(),
                        upR0 = as.numeric())
  
  
  # For every possible simulation
  for(c in 1:(ncol(df - 1))){
    
    # Get incidence
    incidence <- data.frame(cases = df[,c])
    incidence$cases <- rbinom(n = length(incidence$cases),
                              size = incidence$cases,
                              prob = rho)
    incidence$day <- 2*7 + 1:dim(incidence)[1]
    
    # Make sure there are cases in the first 2 observed weeks...
    somecases <- (sum(incidence$cases[1:14]) >= 12)
    
    if (somecases){
      #== simple EpiEstim: biweekly sliding window
      
      epestim <- epiestim(incidence[1:14, 1:2], mean_GT = mean_si, sd_GT = sd_si)
    
      results <- rbind(results, data.frame(
        sim = c, method = 'simpleEpiEstim', R0 = epestim[1],
        lowR0 = epestim[2], upR0 = epestim[3]
      ))
      
      
      #== Linear Exponential Growth: estimate using first 2 weeks 
      if(any(incidence$cases==0)){
        incidence$casesC <- incidence$cases+0.5
        incidence$logcases <- log(incidence$casesC)
      }else{
        incidence$logcases <- log(incidence$cases)
      }
      
      eglin <- EG_Lin(data = incidence[1:14, 1:2], mean_GT = mean_si, sd_GT = sd_si)
      
      results <- rbind(results, data.frame(
        sim = c, method = 'EgLin', R0 = eglin[1],
        lowR0 = eglin[2], upR0 = eglin[3]
      ))
      
      
     
      #== EpiEstim with correction

      epestim2 <- epiestim_adjust(data = incidence[1:14, 1:2], mean_GT = mean_si, sd_GT = sd_si, adjust = TRUE)
      
      epestim2 <- epestim2$results
      results <- rbind(results, data.frame(
          sim = c, method = 'EpiEstim_adjusted', R0 = epestim2[1],
          lowR0 = epestim2[2], upR0 = epestim2[3]
      ))
      
    }
  }
    
    
  rownames(results) <- c()
  
  
  # show results:
  alpha <- gsub('pop=(.*?)$','',name)
  alpha <- as.numeric(gsub('alpha=','',alpha))
  trueR0 = 2*5^{alpha-1}
  
  results$method <- as.factor(plyr::revalue(results$method, 
                                            c("EgLin"="EG", "EpiEstim_adjusted"="EpEsAdj", "simpleEpiEstim"="EpEs")))
  results <- results %>% group_by(method)
  
  # This is stupid but otherwise I can t get it to work:
  if(alpha == 0.7){
    gg <- ggplot(data = results, aes(y = R0, x = method, fill = method)) + geom_violin()  +
      geom_hline(aes(yintercept = 3*5^{0.7-1}),col = "red", linetype = "dashed")
    }
  if(alpha == 0.75){
    gg <- ggplot(data = results, aes(y = R0, x = method, fill = method)) + geom_violin()  +
      geom_hline(aes(yintercept = 3*5^{0.75-1}),col = "red", linetype = "dashed")
  }
  if(alpha == 0.8){
    gg <- ggplot(data = results, aes(y = R0, x = method, fill = method)) + geom_violin()  +
      geom_hline(aes(yintercept = 3*5^{0.8-1}),col = "red", linetype = "dashed")
  }
  if(alpha == 0.85){
    gg <- ggplot(data = results, aes(y = R0, x = method, fill = method)) + geom_violin() +
      geom_hline(aes(yintercept = 3*5^{0.85-1}),col = "red", linetype = "dashed")
  }
  if(alpha == 0.9){
    gg <- ggplot(data = results, aes(y = R0, x = method, fill = method)) + geom_violin() +
      geom_hline(aes(yintercept = 3*5^{0.9-1}),col = "red", linetype = "dashed")
  }
  if(alpha == 0.95){
    gg <- ggplot(data = results, aes(y = R0, x = method, fill = method)) + geom_violin() +
      geom_hline(aes(yintercept = 3*5^{0.95-1}),col = "red", linetype = "dashed")
  }
    if(alpha == 1){
    gg <- ggplot(data = results, aes(y = R0, x = method, fill = method)) + geom_violin() +
      geom_hline(aes(yintercept = 3),col = "red", linetype = "dashed")
  }

  # fix style
  gg <- gg + theme_minimal() +  theme(legend.position = 'none') + labs( x = '', y = 'R0 estimate')
  # common xylimits
  gg <- gg + ylim(c(0,4))
  
  list_plots[[name]] <- gg
  
  results$inside <- (trueR0 <= results$upR0 & trueR0 >= results$lowR0)
  results$CIwidth <- results$upR0 - results$lowR0

  tmp <- results %>% group_by(method) %>%
    summarise(coverage = mean(inside), CIwidth = mean(CIwidth),
              q_Rt05 = quantile(R0, .05), q_Rt50 = quantile(R0, .50), q_Rt95 = quantile(R0, .95),
              R0 = trueR0)
  
  final_stats <- rbind(final_stats, tmp)
  
}
knitr::kable(final_stats)

final_stats$rho <- rho


figure <- ggarrange(plotlist = list_plots, 
                    labels = names(sims),
                    ncol = 2, nrow=3)
figure
```

We note that, the smaller the value of $\alpha$, the more the Exponential Growth method underestimates the basic reproduction number. 
This is to be expected when not observing the first generations of infections.
Due to the sub-exponential growth assumption, the growth rate inferred from data on days $t \in \{15, ..., 28\}$ will already have decreased compared to the growth rate at time $t=0$, yielding to underestimation of $R_0$. 
Consequently, this underestimation yields to underestimation of $R_0$ in our proposed method.
Notice however that EG and EpiEsAdj never overestimate $R_0$.

On the other hand, the difference between  $R_0$ and EpiEstim estimates increases with $\alpha$.
It would seem like there are two sources of bias which balance each other out.
On one hand, missing generations of infections will yield to higher estimates of $R_0$, as discussed in the paper and shown in our analyses. 
On the other hand, the sub-exponential assumption entails that the effective reproductive number after a couple of generations is lower than the basic reproduction number, independently of the proportion of susceptible individuals.

We therefore note that, although our adjustment is not ideal for small values of $\alpha$, it performs well when considering light departures from exponential growth.
Further, a back-imputation procedure similar in spirit to the one we propose could be implemented when the disease is known to spread at a sub-exponential rate.
In the case of a polynomial rate of degree $p$, linear regression could be performed on the $p$-th rooted incidence.
If the degree of the polynomial $p$ is not known in advance, then we could expect the back-imputation procedure to be sensitive to the observations and a longer time window would be needed to jointly estimate $p$ and $r$. 


```{r include=FALSE}
ggsave(file.path(plots_path,paste0('figure2_sens_subexp_',alpha,'.png')),figure, w = 10, h = 11)
ggsave(file.path(plots_path,paste0('figure2_sens_subexp_',alpha,'.pdf')),figure, w = 10, h = 11)
ggsave(file.path(plots_path,paste0('figure2_sens_subexp_',alpha,'.tiff')),figure+reqs, device='tiff', w=10, h=11, dpi=600)
```


