---
title: "Report Simulations"
author: "Andrea Brizzi"
date: "21/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Dependencies
require(EpiEstim)
require(dplyr)
require(ggplot2)
require(png)
require(ggpubr)

# set up directories' paths
repo_path <- '~/Documents/mini_project_2/R0_simulations'
setwd(repo_path)
data_path <- '~/Documents/mini_project_2/R0_simulations/data/'
plots_path <- '~/Documents/mini_project_2/plots'

# sources 
source("methods.R")

# extracts legend
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

MGF_integral <- function(x, mean_GT, sd_GT){
  
  shape <- (mean_GT^2/sd_GT^2)
  scale <- (sd_GT^2/mean_GT)
  out <- c()
  
  for (r in x){
    if (r > -1/scale){
      out <- c(out, (1 - scale * (-r))^(-shape) )
    }else{
      out <- c(out, NA)
    }
  }
  
  return(out)
}
```


In this report, we want to compare different estimates of EpiEstim and Exponential Growth in different scenarios.
We assume a perfect reporting rate $\rho = 100 \%$, and see how estimates evolve when missing 0/2/4 generations of infections, with and without adjustment.

The aim is to 'value' the contribution of our adjustment as well as characterise the biases due to either low numbers of data points, or missing generations.

We start by loading up the data as always.

```{r}
sims_SEIR <- readRDS(file.path(data_path, "sims_SEIR"))
names <- names(sims_SEIR)[c(2,4,6,8)]

# generation interval parameters
mean_si <- 6.5
sd_si <- sqrt(3^2 + 3.5^2)

# reporting rate
rho <- 1
```

Now we move to apply our methods to the simulations.
Should I write a function that does that?
Consider adding a condition which does not allow to estimate when there are too little cases.

```{r}

results <- data.frame(trueR0 = as.numeric(),
                      missedGens = as.numeric(),
                      week_of_estimate = as.numeric(),
                      method = as.character(),
                      estR0 = as.numeric(),
                      lowCI = as.numeric(),
                      upCI = as.numeric())


for (name in names){
  
  # extract true R0
  trueR0 <- as.numeric(stringr::str_match(name, "R0=\\s*(.*?)\\s*pop")[2])
    
  # Get simulations with given R0 and focus on first 6 weeks anyways 
  df <- sims_SEIR[[name]][1:(7*6),]
  
  for (sim in 1:100){
    
    incidence <- df[, sim]
    incidence <- data.frame(cases = incidence, day = seq_along(incidence))
    
    # For every option of missed generations get estimates from first available data to the end
    for (miss_gens in c(0,1,2)){
      
      data <- filter(incidence, day >= miss_gens * 7)
      
      # define 14 days sliding windows on which to estimate.
      # Inside [] we have the function mapping the n of missgens to the number of possible intervals
      # 0 -> 3; 2 -> 2; 4 -> 1
      # 0 -> 3; 1 -> 2; 2 -> 1
      n_intervals <- 3 - miss_gens
      t_start <- c(2, 8, 15)[1:n_intervals] 
      t_end <- c(14, 21, 28)[1:n_intervals] 
      
      # == Apply normal EpiEstim
      epestim <- EpiEstim::estimate_R(c(data$cases), method = "parametric_si",
                           config = make_config(
                             list(
                               mean_si = mean_si,
                               std_si = sd_si,
                               t_start = t_start,
                               t_end = t_end
                             )))$R[,c("Mean(R)", "Quantile.0.025(R)", "Quantile.0.975(R)")]
      
      # put in right format
      names(epestim) <- c('estR0', 'lowCI', 'upCI')
      epestim$week_of_estimate <- (3-n_intervals):2
      epestim$method <- 'EpiEstim'
      
      
      # == Apply EpiEstim adjusted
      
      # extend the data for required number of days.
      tmp <- data[1:14,]
      
      if(any(tmp$cases==0)){
        tmp$casesC <- tmp$cases+0.5
        tmp$logcases <- log(tmp$casesC)
      }else{
        tmp$logcases <- log(tmp$cases)
      }
      
      # EG_growth
      exp_fit <- lm(logcases ~ day, data = tmp)
      1/MGF_integral(exp_fit$coefficients[2], mean_GT = mean_si, sd_GT = sd_si)
      1/MGF_integral(confint(exp_fit, level=0.95)[2, c(1,2)], mean_GT = mean_si, sd_GT = sd_si)
      
      # introduce extrapolated data
      day0 <- data$day[1] - 1
      tmp <- data.frame(day =  (-21):0 + day0,
                        logcases = predict.lm(exp_fit, newdata = data.frame(day = (-21):0  + day0 )))

      if (any(data[1:14,]$cases==0)){
        tmp$casesC <- exp(tmp$logcases)
        tmp$cases <- tmp$casesC - 0.5
        tmp$cases[tmp$cases < 0] <- 0
      }else{
        tmp$cases <- exp(tmp$logcases)
      }
    
      data <- rbind(tmp[,c('cases', 'day')], data)
      
      # adjust windows 
      t_start[1] <- 1
      t_start <- t_start + dim(tmp)[1]
      t_end <- t_end + dim(tmp)[1]
      
      epestim_adj <- EpiEstim::estimate_R(c(data$cases), method = "parametric_si",
                           config = make_config(
                             list(
                               mean_si = mean_si,
                               std_si = sd_si,
                               t_start = t_start,
                               t_end = t_end
                             )))$R[,c("Mean(R)", "Quantile.0.025(R)", "Quantile.0.975(R)")]

      # put in right format
      names(epestim_adj) <- c('estR0', 'lowCI', 'upCI')
      epestim_adj$week_of_estimate <- (3-n_intervals):2
      epestim_adj$method <- 'EpiEstim_adj'
      
      tmp <- rbind(epestim, epestim_adj)
      tmp$trueR0 <- trueR0
      tmp$missedGens <- miss_gens
      
      results <- rbind(results, tmp)
    }
    
  }
} 
```

Now that we have the results, we can analyze them. Let us focus with a relatively high R0.

```{r}
R0VALUE <- 2.5

analysis <- filter(results, trueR0 == R0VALUE)
analysis <- analysis %>% group_by(missedGens, week_of_estimate, method)
analysis$week_of_estimate <- as.factor(analysis$week_of_estimate)
analysis$week_of_estimate <- as.factor(plyr::revalue(analysis$week_of_estimate,
                                                     c("0"="0-2", "1"="1-3", "2"="2-4")))

analysis$missedGens <- as.factor(analysis$missedGens)

analysis$method <- gsub('EpiEstim_adj', 'EpEsAd', analysis$method)
analysis$method <- gsub('EpiEstim', 'EpEs', analysis$method)
gg <- ggplot(filter(analysis), aes(x = week_of_estimate, y = estR0, fill = method, color = method))  +
  geom_violin() + 
  geom_hline(aes(yintercept = R0VALUE), linetype = 'dashed') + 
  facet_grid(rows = vars(missedGens)) + 
  theme_bw() + xlab('window used') + ylab(expression("R"[0])) + theme(legend.position = 'bottom')
gg
ggsave(file.path(plots_path, 'figure4.png'), w = 8, h = 6.5)
a <- analysis %>%  summarise( bias = mean(estR0) - trueR0,
                              CIwidth = mean(upCI - lowCI),
                              coverage = mean(trueR0 <= upCI & trueR0 >= lowCI))

a <- unique(a)


```



